State of the Practice for Lattice Boltzmann Method Software

### 2021-12-29 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Replied to LBM paper comments re traceability and understandability

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\zm}[1]{\authornote{magenta}{MN}{#1}} %Zahra Motamed
\newcommand{\pmm}[1]{\authornote{cyan}{AD}{#1}} %Peter Michalski

@@

Software developers noted that they believe users have generally found their
packages to be understandable. \wss{Confusion between understandability and
usability - threat to validity example.}  The interviewed developer of ESPResSo
commented that some users have attempted to run physically incompatible LBM
methods, and the solution was to edit the code to prevent such combinations, as
well as to update the documentation to prevent misunderstanding the methods.
Similarly, a developer of pyLBM noted that some users had issues setting up
parameters for LBM schemes. The solution to this was to update the interface
where these parameters are set, as well as to add functionality to test the
stability of the parameters. A developer of OpenLB noted that some users lack
the background knowledge to easily model fluid dynamics problems using their
software. A frequently asked questions section was added to their package
website to help users find answers to common questions. The package also has
detailed documentation, including guides and usage requirements specification,
to better help users understand the software.  \wss{Peter - the discussion here
is more about user usability than the understandability of the code for
developers.  Did any of the developers discuss developer understandability?  The
confusion between understandability and usability was also a confusion during
the interviews Ao conducted.}

@@

The interviewed developer of ESPResSo noted that all major additions to their
package had accompanying changes to artifacts and documentation. They noted that
considerable effort had been put into the documentation. They further commented
that they want to lower the entry barrier for new developers, and because of
that their package has a considerable amount of developer documentation. This
documentation informs developers on how to get started, and orients them to the
artifacts, source code, and system architecture, as well as how the software
package build system works, and how the coupling between the simulation engine
and the interface works.  \wss{Peter - these points about developer
documentation seem like they would fit better under understandability, don't you
think?}  \wss{If ESPresSo scored high (I think it did), we can use that to
further promote the value of documentation.  Also a threat to validity is that
we have implicitly assumed documentation is valuable when measuring quality.}

Developers noted the importance of documentation for both the users and
developers of their software. New features are always added to the
documentation. The developers use documentation to stay up to date on the status
of the software package, and to help expand features, like computational models
or algorithms. This is necessary so that the coding standard for these models is
kept consistent for new developers. \wss{Peter - this also sounds like
discussion that is relevant for understandability.}

The importance of documentation for both users and developers was stressed
throughout the interviews. However, it was noted several times that a lack of
time and funding has a negative affect on the documentation. Most of the
developers are scientific researchers evaluated on the scientific papers that
they produce. Writing and updating documentation is something that is done in
their free time, if that time arises. Sometimes it is a last priority for the
developers. Finding ways to hasten updating documentation would increase the
frequency of such updates and benefit both users and developers.  \wss{Same
thing has come up in papers.  I'll try to find a reference for this.  This is
also something we can point to for recommendations related to code generation.}
\wss{Peter - this also sounds like understandability related discussion?  I
wonder if we should drop Traceability as a quality that we assessed?  There
doesn't seem to be much content specific to traceability.}  The developer of
OpenLB noted the use of documentation generators like Doxygen. It would be
advisable for more projects to use such automatic document generation tools,
since some projects do not do this.  \wss{We can refer back to this for the
recommendations.}

@@

For the developers interviewed, visibility is not currently a priority.
Visibility is not particularly necessary because the teams are fairly small.
Although they accept outside contributors, these are rare and the teams remain
tight-knit.  Most of the teams interviewed have all members at the same
institution, although one team is spread out internationally. \wss{Peter, can
you say which team is international, so that we can add this detail to the
paper?}  For teams of only a few developers, new features or major changes are
discussed with the entire team. Projects with more than a few developers have
lead developer roles. These lead developers review potential additions to the
software.

@@

We analyzed the state of the practice of software development in the Lattice
Boltzmann Methods software domain by quantitatively and qualitatively measuring,
and comparing, 23 software packages along quality attributes. The software
qualities that were assessed in this report are listed in Section
~\ref{softwarequalities}. A methodology for assessing the state of the practice
of software development in SCS domains was presented. Domain packages were
assessed to answer the software development related research questions listed in
Section~\ref{secIntro} to understand how software quality is impacted by software
development choices, including principles, processes, and tools. Software
developers were interviewed to identify development pain points, and to identify
how software quality is ensured. Quantitative data was used to rank the software
packages using the AHP. The ranking designations were compared with rankings
from the software development community, and we found that many of our top 10
ranked packages are ranked highly by a domain expert. Recommendations for
improving software along quality metrics were made, and highlights are presented
in Section~\ref{highlightedrecommendations} of this conclusion. The findings of
this report can be used to guide future development of SCS, specifically along
quality attributes, and to reduce software quality failures.

>>>>

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\zm}[1]{\authornote{magenta}{MN}{#1}} %Zahra Motamed
\newcommand{\pmm}[1]{\authornote{cyan}{PM}{#1}} %Peter Michalski

@@

Software developers noted that they believe users have generally found their
packages to be understandable. \wss{Confusion between understandability and
usability - threat to validity example.}  The interviewed developer of ESPResSo
commented that some users have attempted to run physically incompatible LBM
methods, and the solution was to edit the code to prevent such combinations, as
well as to update the documentation to prevent misunderstanding the methods.
Similarly, a developer of pyLBM noted that some users had issues setting up
parameters for LBM schemes. The solution to this was to update the interface
where these parameters are set, as well as to add functionality to test the
stability of the parameters. A developer of OpenLB noted that some users lack
the background knowledge to easily model fluid dynamics problems using their
software. A frequently asked questions section was added to their package
website to help users find answers to common questions. The package also has
detailed documentation, including guides and usage requirements specification,
to better help users understand the software. \wss{Peter - the discussion here
is more about user usability than the understandability of the code for
developers. Did any of the developers discuss developer understandability?  The
confusion between understandability and usability was also a confusion during
the interviews Ao conducted.} \pmm{The developer of ESPResSo did comment on developer understandability, specifically for developers that are/were part of their peer review process. He commented that code architecture is documented, and that code is made to be readable, and that these help peer reviews of the code.}

@@

The interviewed developer of ESPResSo noted that all major additions to their
package had accompanying changes to artifacts and documentation. They noted that
considerable effort had been put into the documentation. They further commented
that they want to lower the entry barrier for new developers, and because of
that their package has a considerable amount of developer documentation. This
documentation informs developers on how to get started, and orients them to the
artifacts, source code, and system architecture, as well as how the software
package build system works, and how the coupling between the simulation engine
and the interface works.  \wss{Peter - these points about developer
documentation seem like they would fit better under understandability, don't you
think?}  \pmm{After reviewing the definitions of the two qualities, I think you are right. I think that the point presented in the first sentence of this paragraph belongs here in traceability (but could be reworded to note that ESPResSo also has detailed change logs in their docs - change logs are also available for Palabos and waLBerla), but the rest of the paragraph can be moved to understandability.} \wss{If ESPresSo scored high (I think it did), we can use that to
further promote the value of documentation.  Also a threat to validity is that
we have implicitly assumed documentation is valuable when measuring quality.}

Developers noted the importance of documentation for both the users and
developers of their software. New features are always added to the
documentation. The developers use documentation to stay up to date on the status
of the software package, and to help expand features, like computational models
or algorithms. This is necessary so that the coding standard for these models is
kept consistent for new developers. \wss{Peter - this also sounds like
discussion that is relevant for understandability.} \pmm{The points here sound like they could work for both. The definition of traceability notes that `This quality refers to keeping track of information as it changes forms or relates between artifacts'. So the point `New features are always added to the documentation' could be reworded to fit that part of the definition of traceability. As noted above, this is similar to the point about some packages having including detailed change logs. Same with `The developers use documentation to stay up to date on the status of the software package, and to help expand features'. I agree that at least some of this paragraph could be in understandability, especially the first and second sentences.}

The importance of documentation for both users and developers was stressed
throughout the interviews. However, it was noted several times that a lack of
time and funding has a negative affect on the documentation. Most of the
developers are scientific researchers evaluated on the scientific papers that
they produce. Writing and updating documentation is something that is done in
their free time, if that time arises. Sometimes it is a last priority for the
developers. Finding ways to hasten updating documentation would increase the
frequency of such updates and benefit both users and developers.  \wss{Same
thing has come up in papers.  I'll try to find a reference for this.  This is
also something we can point to for recommendations related to code generation.}
\wss{Peter - this also sounds like understandability related discussion?  I
wonder if we should drop Traceability as a quality that we assessed?  There
doesn't seem to be much content specific to traceability.}  \pmm{Based on the comments I just added above, I think we can keep at least a small part of what is here in traceability, rewording parts to highlight the importance of change logs. Some of the content in this section should indeed be moved to understandability.} The developer of OpenLB noted the use of documentation generators like Doxygen. It would be advisable for more projects to use such automatic document generation tools, since some projects do not do this.  \wss{We can refer back to this for the recommendations.}

@@

For the developers interviewed, visibility is not currently a priority.
Visibility is not particularly necessary because the teams are fairly small.
Although they accept outside contributors, these are rare and the teams remain
tight-knit.  Most of the teams interviewed have all members at the same
institution, although one team is spread out internationally. \wss{Peter, can
you say which team is international, so that we can add this detail to the
paper?} \pmm{The interviewed developer of OpenLB noted that there were tasks for local and international developers when discussing their software development model. He also noted that they have frequent local meetings and annual international meetings.}  For teams of only a few developers, new features or major changes are
discussed with the entire team. Projects with more than a few developers have
lead developer roles. These lead developers review potential additions to the
software.

@@

We analyzed the state of the practice of software development in the Lattice
Boltzmann Methods software domain by quantitatively and qualitatively measuring,
and comparing, 24 software packages along quality attributes. The software
qualities that were assessed in this report are listed in Section
~\ref{softwarequalities}. A methodology for assessing the state of the practice
of software development in SCS domains was presented. Domain packages were
assessed to answer the software development related research questions listed in
Section~\ref{secIntro} to understand how software quality is impacted by software
development choices, including principles, processes, and tools. Software
developers were interviewed to identify development pain points, and to identify
how software quality is ensured. Quantitative data was used to rank the software
packages using the AHP. The ranking designations were compared with rankings
from the software development community, and we found that many of our top 10
ranked packages are ranked highly by a domain expert. Recommendations for
improving software along quality metrics were made, and highlights are presented
in Section~\ref{highlightedrecommendations} of this conclusion. The findings of
this report can be used to guide future development of SCS, specifically along
quality attributes, and to reduce software quality failures.
